# Machine learning

# Setup

```{r setup4, echo=FALSE}
# setting up general settings for chunks
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE
)
```

```{r loading libraries, warning=TRUE}
# loading libraries
library(here)
library(tidyverse)
library(reactable)
```

### Plan of action

##### Introduction

During the "Data Science for biology 2" course every student has the opportunity to learn a new skill for their portfolio. I (Thijmen Weijgertze) have decided to learn the myself the basics of "Machine Learning". This skill will be learned in a span of ca. 4 weeks (around 30 hours). To learn this skill I have set up a goal with a corresponding planning to achieve the this goal.

##### Goal

At the end of this course I want to have created an algorithm able to predict the recurrence of breast cancer using the breast-cancer.data dataset. This algorithm will be written in R using principles from "tidymodels", "CART", "random forest" and possibly other Machine Learning techniques I'll come across.

##### Planning

- Week 1 (7th - 14th of September): Delving into the principles of "tidymodels", "CART", "random forest"
- Week 2 (14th - 21th of September): Writing my first Machine Learning algorithm
- Week 3-4 (21th - 6th of June): Apply the learned skills on the breast-cancer.data dataset to predict the recurrence of breast cancer

##### Dataset

```{r loading in data}
# loading in data
bcData <- read.csv( 
  here::here(
    "data_machineLearning",
    "breast-cancer.data"
  ), 
  header = FALSE, # dataset does not contain column names
  col.names = c("class", "age", "menopause", "tumor_size", "inv_nodes", "node_caps", "deg_malig", "breast", "breast_quad", "irradiat") # setting column names
)

# presenting data in table form
reactable(
  bcData, 
  filterable = TRUE,
  compact = TRUE,
  bordered = TRUE,
  defaultPageSize = 5
  )
```

This breast cancer domain was obtained from the University Medical Centre, Institute of Oncology, Ljubljana, Yugoslavia. Thanks go to M. Zwitter and M. Soklic for providing the data. The breast-cancer.data domain consist of a total of `r nrow(bcData)` rows and `r ncol(bcData)` columns. From which `r bcData %>% filter(class == "no-recurrence-events") %>% nrow()` instances are no-recurrence-events and `r bcData %>% filter(class == "recurrence-events") %>% nrow()` instances are recurrence-events. 

## The Tidymodels get-started-guide

In order to build a model the tidymodels package and some additional packages must be installed.

```{r installing packages for model building, eval=FALSE}
# install.packages("tidymodels") # main package
# 
# # helper packages in order to use the data included in the tidymodels guide
# install.packages("readr")
# install.packages("broom.mixed")
# install.packages("dotwhisker")
```

```{r loading packages for model building}
# loading in the installed packages
library(tidymodels)
library(readr)
library(broom.mixed)
library(dotwhisker)
```

For this guide sea urchin data can be loaded from https://tidymodels.org/start/models/urchins.csv

```{r}
# loading in the data 
urchins <- read.csv("https://tidymodels.org/start/models/urchins.csv")

# inspecting the data and data types
reactable(
  urchins, 
  filterable = TRUE,
  compact = TRUE,
  bordered = TRUE,
  defaultPageSize = 5
  )
```


