---
title: "C. Elegans plate experiment"
author: "Thijmen Weijgertze"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: lumen
---
<div style="text-align: justify">

```{r setup, echo=FALSE}
# setting up general settings for chunks
knitr::opts_chunk$set(
  warning=FALSE,
  message=FALSE
)
```

### **Setup**

##### Setting a seed and loading packages

```{r setting seed}
# Seed chosen based on the current year
set.seed(2023)
```

```{r packages, message=FALSE, warning=FALSE}
# Loading packages
library(tidyverse)
library(readxl)
library(here)
library(reactable)
library(RColorBrewer)
```

##### Importing and inspecting the data

```{r importing and inspecting the data}
# importing xlsx file
CE_LIQ_FLOW_062_Tidydata <- read_excel(
  here::here(
    "data-raw",
    "CE.LIQ.FLOW.062_Tidydata.xlsx"
  )
)

# inspecting data in table format with the reactable package
reactable(
  CE_LIQ_FLOW_062_Tidydata, 
  filterable = TRUE,
  compact = TRUE,
  bordered = TRUE,
  defaultPageSize = 5
  )
```

### **Scatterplot**

##### Pseudocode 

- Deciding which columns will be included in the scatterplot
- Checking and possibly changing the datatypes of those columns 
- Retrieving log10(0) values by filtering on compConcentration == 0
- Plotting the data in a scatterplot using ggplot
- Setting the x-axis to a log10 scale 
- Adding jitter to spread out points on top of eachother

##### Checking and correcting datatypes needed for the scatterplot

```{r checking and correcting datatypes}
# Checking the data types of the following columns: RawData, compName, expType and compConcentration
CE_LIQ_FLOW_062_Tidydata %>% select(RawData, compName, compConcentration, expType) %>% str()

# Changing compConcentration to numeric; changing compName into a factor; changing expType into a factor
CE_LIQ_FLOW_062_Tidydata$compConcentration <- parse_number(CE_LIQ_FLOW_062_Tidydata$compConcentration)
CE_LIQ_FLOW_062_Tidydata$compName <- factor(CE_LIQ_FLOW_062_Tidydata$compName, levels = unique(CE_LIQ_FLOW_062_Tidydata$compName))
CE_LIQ_FLOW_062_Tidydata$expType <- factor(CE_LIQ_FLOW_062_Tidydata$expType, levels = unique(CE_LIQ_FLOW_062_Tidydata$expType))

# Checking the new dataypes and factor levels
CE_LIQ_FLOW_062_Tidydata %>% select(RawData, compName, compConcentration, expType) %>% str()
levels(CE_LIQ_FLOW_062_Tidydata$compName)
levels(CE_LIQ_FLOW_062_Tidydata$expType)
```

##### Table with log10(0) values per level in expType

```{r table with log10(0) values}
# Count the total of rows per level in expType
CE_LIQ_FLOW_062_Total <- 
  CE_LIQ_FLOW_062_Tidydata %>% 
  group_by(expType) %>%
  summarise(total_rows = length(expType))

# Count rows with a log10 equal to 0 per level in expType
CE_LIQ_FLOW_062_filtered <- 
  CE_LIQ_FLOW_062_Tidydata %>% 
  filter(compConcentration == 0) %>% 
  group_by(expType) %>%
  summarise(log10_of_0_rows = length(expType))

# Joining tibbles
left_join(CE_LIQ_FLOW_062_Total, CE_LIQ_FLOW_062_filtered, by="expType")

# Counting all log 10 values to add them in the side note of the scatterplot 
log_10_0_sum <- sum(CE_LIQ_FLOW_062_filtered[, 'log10_of_0_rows'], na.rm = TRUE)

# https://stackoverflow.com/questions/9676212/how-to-sum-data-frame-column-values
# https://stackoverflow.com/questions/26114525/how-to-count-how-many-values-per-level-in-a-given-factor#:~:text=Or%20using%20the%20dplyr%20library%3A%20library%20%28dplyr%29%20set.seed,similar%20to%20the%20use%20of%20pipes%20in%20bash.
```




##### Plotting the CE_LIQ_FLOW_062_Tidydata in a scatterplot

```{r scatterplot, fig.cap="The positive control for this experiments is controlPositive\nThe negative control for this experiment is controlNegative"}
# Plotting the CE_LIQ_FLOW_062_Tidydata in a scatterplot
ggplot(
    data = CE_LIQ_FLOW_062_Tidydata, 
    aes(x = log10(compConcentration), y = RawData)
    )+
  geom_point(
    aes(color = compName, shape = expType),
    size = 1.5, 
    alpha = 0.8,
    position = position_jitter(width = 0.1) # jitter to spread out values on top of eachother
    )+
  labs(
    title = "Scatterplot CE_LIQ_FLOW_062_Tidydata",
    caption = paste("Side note:", log_10_0_sum, "values containing a log10 of 0 are not visible in the graph"),
    x = "log10(compConcentration) in nM"
  )+
  scale_color_brewer(palette = "Dark2") # colorblind friendly color scale
```

### **Statistical tests**

##### Pseudocode 

- 

</div> 