# Relational databases

```{r setup8, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo=TRUE)
```

### Setup

```{r loading relational database packages}
# loading packages
library(tidyverse)
library(dslabs)
library(here)
library(reactable)
library(twPackage)
```

The flu and dengue data sets are retrieved from Google Inc. [*@googleinc.GoogleFluTrends2015 ; @googleinc.GoogleDengueTrends2015*]. The gapminder data set is retrieved from the dslabs package [*@irizarryDslabsDataScience2023*].

### Reading in files

```{r}
# loading flu data
flu_data <- read.csv(
  here::here(
    "data_relationalDatabases",
    "flu_data.csv"
  ),
  skip = 11 # skip the first 11 lines
)
# loading dengue data
dengue_data <- read.csv(
  here::here(
    "data_relationalDatabases",
    "dengue_data.csv"
  ),
  skip = 11 # skip the first 11 lines
)
# loading gapminder data
gap_data <- gapminder
```

### Inspecting data

```{r}
# inspecting flu data
tw_table(flu_data, pagesize = 5)

# inspecting dengue data 
tw_table(dengue_data, pagesize = 5)

# inspecting gapminder data 
tw_table(gap_data, pagesize = 5)
```

The flu dataset must be changed into tidy format. The flu and dengue data tidy the countries will be stored under one column named "country". The date column will be separated into the columns: year, month and day.

### Making the data tidy

```{r}
## making flu data tidy
# creating the country column
flu_tidy <- pivot_longer(
  data = flu_data,
  cols = c(Argentina:Uruguay),
  values_to = "search_activity",
  names_to = "country"
    )
# creating year month and date columns
flu_tidy <- separate(
  flu_tidy,  
  Date,  
  into = c("year", "month", "day"), 
  convert = TRUE, 
  sep = "-"
  )

## making dengue data tidy
# creating the country column
dengue_tidy <- pivot_longer(
  data = dengue_data,
  cols = c(Argentina:Venezuela),
  values_to = "search_activity",
  names_to = "country"
    )
# creating year month and date columns
dengue_tidy <- separate(
  dengue_tidy,  
  Date,  
  into = c("year", "month", "day"), 
  convert = TRUE, 
  sep = "-"
  )

# the gapminder data is already tidy
gap_tidy <- gap_data

```

### Inspecting the tidy data

```{r}
# inspecting flu tidy
tw_table(flu_tidy, pagesize = 5)

# inspecting dengue tidy 
tw_table(flu_tidy, pagesize = 5)

# inspecting gapminder tidy 
tw_table(flu_tidy, pagesize = 5)
```

### Exporting the tidy data as .csv and .rds format

```{r, eval=F}
# exporting flu tidy to .csv and .rds format
tw_csv_rds(flu_tidy, "data_relationalDatabases/flu_tidy")

# exporting dengue tidy to .csv and .rds format
tw_csv_rds(dengue_tidy, "data_relationalDatabases/dengue_tidy")

# exporting gap tidy to .csv and .rds format
tw_csv_rds(gap_tidy, "data_relationalDatabases/gap_tidy")
```

### Making connection to the DBeaver database

```{r loading packages to connect R with database}
# loading packages to connect R with the database
library(DBI)
```

```{r connecting with database}
# connecting to database
con <- dbConnect(RPostgres::Postgres(), 
                 dbname = "workflowsdb", 
                 host="localhost", 
                 port="5432", 
                 user="postgres", 
                 password=readRDS(here::here("db_pass.rds"))) # put here your own password
```

### Loading files into the database

```{r loading files into database, eval=F}
# Connect with the database and export the data to DBeaver
dbWriteTable(con, "flu_tidy", flu_tidy)
dbWriteTable(con, "dengue_tidy", dengue_tidy)
dbWriteTable(con, "gap_tidy", gap_tidy)
```

### Inspecting data with SQL

In the following code blocks there will be examples of using SQL to inspect the database data.

```{sql, connection=con}
-- In specting dengue data with SQL
select country from dengue_tidy; 
```

```{sql, connection=con}
-- flu data from the Netherlands without NA values ordered by search_activity
select * 
from flu_tidy 
where 
  search_activity is not null and 
  country='Netherlands'
order by 
  search_activity desc;
```

```{sql ,connection=con}
-- amount of rows in dengue data without NA
SELECT COUNT(*)
FROM dengue_tidy
where search_activity is not null;
```

### joining the tables into a new table

The tables will be joined into a new table. This new table will be used later on to plot the flu_search against the data per country within the South American region. 

```{sql,connection=con, eval=FALSE}
-- joining the tables using SQL
create table joined_tidy as
SELECT -- columns to be included in the table
flu_tidy.year,
flu_tidy.month,
flu_tidy.day,
gap_tidy.continent,
gap_tidy.region,
flu_tidy.country,
flu_tidy.search_activity as flu_search,
dengue_tidy.search_activity as dengue_search
FROM flu_tidy 
  LEFT JOIN dengue_tidy ON -- joining the dengue dataset with the flu dataset based on date and country
    dengue_tidy.year = flu_tidy.year AND
    dengue_tidy.month = flu_tidy.month AND
    dengue_tidy.day = flu_tidy.day AND
    dengue_tidy.country = flu_tidy.country
  LEFT JOIN gap_tidy ON -- joining the gapminder dataset with the flu dataset based on year and country
    gap_tidy.year = flu_tidy.year AND
    gap_tidy.country = flu_tidy.country;
``` 
[*@rahulAnswerCanJoin2014; @SQLJoinInner2016*]

### Inspecting the joined tidy dataset

```{sql connection=con, output.var = "joined_tidy"}
-- inspecting the new table and saving to
select *
from joined_tidy 
```

The data is saved as a variable called joined_tidy by adding 'output.var="joined_tidy"' to the chunk header [@langtangAnswerSQLChunks2022].

```{r}
# inspecting joined tidy variable
tw_table(joined_tidy, pagesize = 5)
```

```{r joined tidy to csv and rds, eval=FALSE}
# exporting joined tidy to .csv and .rds format
tw_csv_rds(joined_tidy, path = "data_relationalDatabases/joined_tidy")
```

```{r disconnecting database}
# disconnecting from the database 
# since the joined_tidy dataset is now stored in a R variable
dbDisconnect(con)
```

### plotting the flu_search against the data per country within the South American region

```{r}
# plotting flu search against the date per South American country
joined_tidy %>% 
  mutate(date = mdy(paste(month,day,year, sep = "-"))) %>% 
  filter(region == "South America") %>% 
  select(flu_search, date, country, region) %>% 
  na.omit() %>% 
  ggplot(aes(x = date, y = flu_search, fill = country)) + 
  geom_col(aes(colour = country)) + 
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1), legend.position = "none") +
  facet_wrap(~country, scales='free')
```
